---
title: "IE6600_Sec03_Group08_Hackathon"
author: "Group 08"
date: "3/2/2021"
output: pdf_document
fig_width: 10 
fig_height: 6
---

  
                                          1.1 Overview
The UNICEF's Dataset contains the key issues affecting the State of the World's Children. We are going to assess the situation of children and women in the areas of education, health, gender equality, rights and protection in form of visualization using R software and try to gain some insight from those visualization to compare the situation of the children and women in the world right now based on different factors which affect them the most.

                                      1.2 Problem statement
Our objective is to find the disparities in school enrollment across various regions, compare child labor and violence discipline with literacy rates achieved . In this report, the main objective is split into 3 sections.
First Section identifies Disparities in school enrollment at various levels and identifies enrollment based on gender. Second section talks on how child labour is associated with lower secondary school enrollment and gender. Section 3  visualizes violent discipline and its relation with literacy.

```{r}
knitr::opts_chunk$set(warning = FALSE,message = FALSE)
```
 
```{r Loading Libraries}
library(gridExtra)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(reshape2)
library(extrafont)
library(tidyr)
library(data.table)
library(scales)
library(readxl)
library(stringr)
library(ggalluvial)
library(ggrepel)
```

```{r Loading the Dataset}
Data <- read_excel("Hackathon_Unicef_Dataset.xlsx", sheet = "Data" )
Data<-transform(Data, Value=as.numeric(Value))
```
# Section 1
```{r}
req_indicators <- Data %>% 
  filter(Data$Indicator == ("Pre-primary school participation - Gross enrolment ratio (%), male") | 
  Data$Indicator ==("Pre-primary school participation - Gross enrolment ratio (%), female")| 
           Data$Indicator==("Primary school participation - Net enrolment ratio (%), male") | 
           Data$Indicator == ("Primary school participation - Net enrolment ratio (%), female") |
           Data$Indicator ==("Lower secondary school participation - Net enrolment ratio (%), male") |
           Data$Indicator == ("Lower secondary school participation - Net enrolment ratio (%), female") |  Data$Indicator == ("Primary school participation - Out-of-school rate of children of primary school age (%), male")|
           Data$Indicator==("Primary school participation - Out-of-school rate of children of primary school age (%), female")|
           Data$Indicator==("Primary school net attendance ratio, Urban") |
           Data$Indicator==("Primary school net attendance ratio, Rural")
         )
req_regions <- req_indicators %>%
  filter((Countries.and.areas=='Central African Republic') | 
        (Countries.and.areas=='East Asia and Pacific')  |
        (Countries.and.areas=='Europe and Central Asia') |
        (Countries.and.areas=='Eastern Europe and Central Asia') |
        (Countries.and.areas=='Western Europe') | 
        (Countries.and.areas=='Middle East and North Africa') |
        (Countries.and.areas=='Eastern and Southern Africa') | 
        (Countries.and.areas=='West and Central Africa')
          )
pivot_Data <- req_regions %>%
  select(Countries.and.areas,Indicator, Value) %>%
  pivot_wider(names_from = Indicator, values_from = Value )
# Selecting required columns
req_df <- Data %>% 
  filter(Data$Indicator == ("Total adult literacy rate (%)") | Data$Indicator ==("Primary school net enrolment ratio (%)")| 
           Data$Indicator==("Pre-primary school participation - Gross enrolment ratio (%), male") | 
           Data$Indicator == ("Pre-primary school participation - Gross enrolment ratio (%), female") |
           Data$Indicator ==("Primary school participation - Gross enrolment ratio (%), male") |
           Data$Indicator == ("Primary school participation - Gross enrolment ratio (%), female") |  
           Data$Indicator == ("Primary school participation - Out-of-school rate of children of primary school age (%), male") |
           Data$Indicator==("Primary school participation - Out-of-school rate of children of primary school age (%), female") |
           Data$Indicator == ("Primary school net enrolment ratio (%)")
         )
# Pivoting the Data on Indicators
p_Data <- req_df %>%
  select(Countries.and.areas,Indicator, Value) %>%
  pivot_wider(names_from = Indicator, values_from = Value )
p_Data <- p_Data %>%
  rename(out_of_school_male=`Primary school participation - Out-of-school rate of children of primary school age (%), male`) %>%
  rename(out_of_school_female=`Primary school participation - Out-of-school rate of children of primary school age (%), female`)
```
Average adult literacy percentage of all the countries is 80. Most of the countries achieved this literacy. Yet, almost 40% of the countries are still under the average literacy rate of 80. In order to improve the rate of literacy, children must be ecourages to enroll in school from primary level. 
```{r}
p_Data$newcol<-0
colnames(p_Data)[15] <- "primary_school_net_enrollment_ratio"
p_Data$primary_school_net_enrollment_ratio <- cut(p_Data$`Primary school net enrolment ratio (%)`, c(0,10,20,30,40,50,60,70,80,90, 100), c("0-10", "10-20", "20-30","30-40", "40-50", "50-60", "60-70", "70-80", "80-90", "90-100" ))
d4<-p_Data %>%
  group_by(primary_school_net_enrollment_ratio) %>%
  summarise(number_of_countries= n()) %>%
  drop_na()
chart_6 <-d4 %>%
  ggplot(aes(x =primary_school_net_enrollment_ratio , y = number_of_countries)) + 
    geom_bar(stat = 'identity', fill='red3', alpha=0.6) + theme(axis.text.x = element_text(angle=0, hjust = 1)) + geom_text(aes(label=number_of_countries),vjust=-0.5,  color="red3",  position = position_dodge(1), size=4)+
    labs(y="number of countries", x = "Net enrollment ratio") +labs(title= "Primary school enrollment ratios across various countries")+ theme(legend.position = "bottom")
chart_6
```
From the above chart, we see that in the first 4 countries, the percentage of children enrolled in primary school is less than 50 per cent. These countries belong to African sub continent
```{r}
d<- mutate(pivot_Data, pre_primary = `Pre-primary school participation - Gross enrolment ratio (%), male` +
             `Pre-primary school participation - Gross enrolment ratio (%), female`)
d1<- mutate(d, Primary = `Primary school participation - Net enrolment ratio (%), male` +
             `Primary school participation - Net enrolment ratio (%), female`)
d2<- mutate(d1, Lower_secondary = `Lower secondary school participation - Net enrolment ratio (%), male` +
             `Lower secondary school participation - Net enrolment ratio (%), female`)

chart_2<- d2 %>%
  select(Countries.and.areas, pre_primary, Primary, Lower_secondary) %>%
  drop_na() %>%
  gather(level, ratio, 2:4) %>%
  ggplot(aes(x = reorder(Countries.and.areas, -ratio) , y = ratio , fill = level)) + 
    geom_bar(stat = 'identity', alpha=0.7) + theme(axis.text.x = element_text(angle=45, hjust = 1)) + theme(legend.position = "top")+
    labs(y="ratio", x = "region") +labs(title= "Pre-Primary,Primary,Lower-Secondary School enrollment") + scale_fill_brewer(palette = "Set1") +  theme(legend.justification = "top")
chart_2
```

Ratio of children enrolling in Lower Secondary school is less in regions involving African countries. In East and Central Africa, number of children enrolling in secondary school after primary level is less compared to other regions. On the other hand, in European, Asian regions, ratios of primary and secondary are nearly equal.

```{r}
df1<-p_Data %>%
  select(Countries.and.areas, out_of_school_male, out_of_school_female)
df1<- mutate(df1, out_of_school_ratio_female_to_male= out_of_school_female/out_of_school_male )
index1<- which(df1$out_of_school_ratio_female_to_male>=2 )
df2 <- df1[index1, c(1:4)]
chart_1<- df2 %>%
  select(Countries.and.areas, out_of_school_male, out_of_school_female) %>%
  drop_na() %>%
  arrange(desc(out_of_school_female)) %>%
  top_n(20, out_of_school_female) %>%
  gather(perc, n, 2:3) %>%
  ggplot(aes(x = reorder(Countries.and.areas, -n) , y = n , fill = perc)) + 
    geom_bar(stat = 'identity', position = 'dodge', alpha=0.7) + theme(axis.text.x = element_text(angle=45, hjust = 1)) +
    labs(y="perecentage", x = "Country", color="Gender") +labs(title= "Highligting differences in Out of school percentages - female vs male")+ theme(legend.position = "top")+ labs(title= "Highligting differences in Out of school percentages - female vs male")+ theme(legend.position = "top")+ scale_fill_brewer(palette = "Set1")
chart_1
```
Least developed countries have a high percentages of children not attending schools. If we compare enrollment of male and females in school, more females are out of schools when compared to males. These differences are obvious in the above countries. In these countries, out of 3 children, there at least 2 girls that are not attending school. This stands true not just for less developed nations but also to highly developed ones such as UK and Norway. Though the out-of-school percentages are really low in these nations, it still holds true.

# Section 2
```{r}
df_partchildm <- Data %>% 
  filter(Data$Indicator ==("Lower secondary school participation - Net attendance ratio (%), male") | Data$Indicator ==("Child labour (%), male"))
df_partchildm1 <- df_partchildm %>%
    dplyr::select("Countries.and.areas",Indicator,Value)
df_partchildm1 <- transform(df_partchildm1, Value = as.integer(Value))
df_partchildm1 <- df_partchildm1 %>%
        pivot_wider(names_from = Indicator, values_from = Value )
partchildm <- ggplot(df_partchildm1, aes(y=`Lower secondary school participation - Net attendance ratio (%), male` ,x=`Child labour (%), male`))  + geom_point(color="darkolivegreen3") +labs(y="Lower secondary school participation, male", x = "Child labour(%), male")
df_partchildf <- Data %>% 
  filter(Data$Indicator ==("Lower secondary school participation - Net attendance ratio (%), female") | Data$Indicator ==("Child labour (%), female"))
df_partchildf1 <- df_partchildf %>%
    dplyr::select("Countries.and.areas",Indicator,Value)
df_partchildf1 <- transform(df_partchildf1, Value = as.integer(Value))
df_partchildf1 <- df_partchildf1 %>%
        pivot_wider(names_from = Indicator, values_from = Value )
partchildf <- ggplot(df_partchildf1, aes(y=`Lower secondary school participation - Net attendance ratio (%), female`,x=`Child labour (%), female`))  + geom_point(color='red3') +labs(y="Lower secondary school participation, female", x = "Child labour(%), female") 
grid.arrange(partchildm,partchildf,ncol = 2) 
```
The above chart is a scatter plot between the "Lower secondary school participation - Net attendance ratio (%)" and "Child labour (%)" for male and female. The plot is approximately same for male and female regardless of their gender.We can conclude from the plot that the country with the highest school participation has the lowest child labour and vice versa.


```{r}
df_childfm <- Data %>% 
  filter(Data$Indicator == ("Child labour (%), female") | Data$Indicator == ("Child labour (%), male"))
df_childfm1 <- df_childfm %>%
    dplyr::select("Countries.and.areas",Indicator,Value)
df_childfm1 <- transform(df_childfm1, Value = as.integer(Value)) %>%
  drop_na()
area_childfm1 <- ggplot(df_childfm1, aes(x = Value, fill=Indicator)) +
           geom_area(stat="bin",alpha=0.7) + scale_fill_brewer(palette="Set1") + labs(y="Number of countries", x="Percentage of child labour")+ggtitle("Child labour across countries")
area_childfm1
```
The height of each colored stack represents the proportion of the child labor gender wise. Child labor percentage varies across a scale of 0 to 60. Most countries have low percentages of child labor. But when compared, female children are more subjected to child labor than male children.

#Section 3 
```{r}
df_discf <- Data %>% 
  filter(Data$Indicator == ("Violent discipline (%), female") | Data$Indicator ==("Youth (15–24 years) literacy rate (%), female") )
df_discf1 <- df_discf %>%
   select(`Countries.and.areas`,Indicator,Value)
df_discf1 <- transform(df_discf1, Value = as.integer(Value))
discf1 <- df_discf1 %>%
        pivot_wider(names_from = Indicator, values_from = Value )
ggdiscf <- ggplot(discf1, aes(y=`Youth (15–24 years) literacy rate (%), female`
,x=`Violent discipline (%), female`))  + geom_point(color="red3") +labs(y="Youth Literacy, female", x = "Child labour(%), male, female")
df_discm <- Data %>% 
  filter(Data$Indicator == ("Violent discipline (%), male") | Data$Indicator ==("Youth (15–24 years) literacy rate (%), male") )
df_discm1 <- df_discm %>%
   select(`Countries.and.areas`,Indicator,Value)
df_discm1 <- transform(df_discm1, Value = as.integer(Value))
discm1 <- df_discm1 %>%
        pivot_wider(names_from = Indicator, values_from = Value )
ggdiscm <- ggplot(discm1, aes(y=`Youth (15–24 years) literacy rate (%), male`
,x=`Violent discipline (%), male`)) + geom_point(color="darkolivegreen3") +labs(y="Youth Literacy, male", x = "Child labour(%), male") 
grid.arrange(ggdiscm,ggdiscf,ncol= 2)
```
Both male and females are subjected to physical punishment or psychological aggression equally. Most countries that adopt violent discipline also happen to be the the countries with good literacy rates.

```{r}

discf2 <- discf1 %>% 
  filter(discf1$`Youth (15–24 years) literacy rate (%), female` >= 95)
discf2 <- discf2 %>% 
  filter(discf2$`Violent discipline (%), female` > 70)
plot <- ggplot(discf2, aes(y=`Youth (15–24 years) literacy rate (%), female`, x=`Violent discipline (%), female`)) + geom_point(color = "red", size = 2) +geom_label_repel(aes(label =`Countries.and.areas`), box.padding=0.35, point.padding = 0.5) + scale_fill_brewer(palette="Set1")  +labs(y="Youth Literacy, female", x = "Violent discipline, female") + ggtitle("Literacy Vs Violent discipline,Female(%)")
plot
```
The above graph is obtained for the violent discipline for female greater than 70% and for youth literacy of female its greater than 95%.Here, we can see that being so high in literacy rate they still have so much of violence for children's.

                                        Conclusion

While the average literacy for countries around the world is 80, 40 percent of the countries are still below this average. These countries are mainly from the African continent. In 4 of these regions, less than 50 percent of the children attend primary schools. Moreover, the number of children continuing in to lower secondary education is less than from primary. Also, we noticed that more females are of out school compared to males. From the section2, take away is that countries with the highest school participation have the lowest child labor. Both males and females in the age group 1-14 are subjected to physical punishment or psychological aggression equally. Countries such as Benin, Chad, Guinea-Bissau, Mali, and Somalia have child labor more than 50(%) for males as well as females. 
Section 3 depicts that most countries that adopt violent discipline also happen to be the countries with good literacy rates. The violent discipline for females is more than 70% despite being the literacy rate of more than 90% in these countries.


